Windowed least-squares model reduction for dynamical systems

### 2019-10-08 main.tex
latest
through space--time, proofing

Researchers have pursued the development of space--time
residual-minimization approaches~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to
address the issues incurred by sequential residual minimization in time.
Space--time residual-minimization approaches compute the space--time
trajectory within a low-dimensional space--time trial subspace (spanned by
a small number of space--time basis vectors) that minimizes the full-order-model
residual over the entire time domain.
We note that space--time residual minimization yields a fundamentlaly
different model than sequential residual minimization in time. In particular, space--time
residual minimization associates with an algebraic system of equations defined
over all space and time, whose solution comprises a \text{fixed} vector of
(space--time) generalized coordinates; on the other hand, sequential residual
minimization in time (often) associates with a system of ODEs whose solution
comprises a \textit{time-dependent} vector of (spatial) generalized
coordinates.
As a result, space--time residual minimization yields models that are equipped with \textit{a priori} error bounds that grow more slowly
in time, and the resulting \textit{trajectory} exhibits a notion of
optimality; this implies that the solution exhibits a notion of optimality
over the entire time domain. 
\EP{This has nothing to do the basis, which was the point I was trying to get across}
However, these techniques also suffer from several
limitations. First, the computational cost of solving the algebraic system arising from space--time
approaches scales cubically with the number of space--time degrees of freedom;
in contrast, the computational cost incurred by standard
spatial-projection-based ROMs is linear in the number of temporal degrees of
freedom, as the attendant solvers can leverage the lower-block triangular
structure of the system arising from the sequential nature of time
evolution. As a result, solving the algebraic systems arising from space--time
projection is generally intractable without applying hyper-reduction in time
\cite{choi_stlspg,constantine_strom}. Second, space--time residual
minimization precludes future
state prediction, as these methods employ space--time basis vectors defined over the
entire time domain of interest, which must have been included in the training
simulations.

@@

Space--time projection methods that employ \spaceTimeAcronym\ trial
spaces~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st,bui_thesis}
aim to overcome the latter two shortcomings of LSPG, which are
shared by all model-reduction methods based on spatial projection followed by
time integration. \EP{This is misleading, going to delete} Because these methods employ \spaceTimeAcronym\ trial
spaces, they reduce both the spatial and temporal dimensions of the full-order
model; further, they yield error bounds that grow more slowly in time and
their trajectories exhibit an optimality property over the entire time domain.

>>>>

Researchers have pursued the development of space--time
residual-minimization approaches~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to
address the issues incurred by sequential residual minimization in time.
Space--time residual-minimization approaches compute the space--time
trajectory within a low-dimensional space--time trial subspace (spanned by
a small number of space--time basis vectors) that minimizes the full-order-model
residual over the entire time domain.
We note that space--time residual minimization yields a fundamentlaly
different model than sequential residual minimization in time. In particular, space--time
residual minimization associates with an algebraic system of equations defined
over all space and time, whose solution comprises a \text{fixed} vector of
(space--time) generalized coordinates; on the other hand, sequential residual
minimization in time (often) associates with a system of ODEs whose solution
comprises a \textit{time-dependent} vector of (spatial) generalized
coordinates.
As a result, space--time residual minimization yields models that are equipped with \textit{a priori} error bounds that grow more slowly
in time, and the resulting \textit{trajectory} exhibits a notion of
optimality; this implies that the solution exhibits a notion of optimality
over the entire time domain. 
\EP{Come back to this, this also has to do with residual minimization}
However, these techniques also suffer from several
limitations. First, the computational cost of solving the algebraic system arising from space--time
approaches scales cubically with the number of space--time degrees of freedom;
in contrast, the computational cost incurred by standard
spatial-projection-based ROMs is linear in the number of temporal degrees of
freedom, as the attendant solvers can leverage the lower-block triangular
structure of the system arising from the sequential nature of time
evolution. As a result, solving the algebraic systems arising from space--time
projection is generally intractable without applying hyper-reduction in time
\cite{choi_stlspg,constantine_strom}. Second, space--time residual
minimization precludes future
state prediction, as these methods employ space--time basis vectors defined over the
entire time domain of interest, which must have been included in the training
simulations.

@@

Space--time projection methods that employ \spaceTimeAcronym\ trial
spaces~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st,bui_thesis}
aim to overcome the latter two shortcomings of LSPG. Because these methods employ \spaceTimeAcronym\ trial
spaces, they reduce both the spatial and temporal dimensions of the full-order
model; further, they yield error bounds that grow more slowly in time and
their trajectories exhibit an optimality property over the entire time domain.

### 2019-10-08 section3/general_form.tex
through space--time, proofing

The \methodAcronym\ framework sequentially computes approximate solutions
$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$, where 
$\approxstateArgnt{n}$ is the solution to the
minimization problem
\begin{equation}\label{eq:tclsrm}
\begin{split}
      &\underset{\statey \in \stspace^n}{\text{minimize}} \; \mathcal{J}^n(\statey), \\
			&\text{subject to } \;  \statey(\timeStartArg{n}) =
\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1}(\timeEndArg{n-1})) & n = 2,\ldots,\nslabs \\
\mathbb{P}^n( \stateFOMIC)& n=1, \end{cases} 
\end{split}
\end{equation}
%\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$ as the solution to the minimization problem,
%\begin{equation}\label{eq:tclsrm}
%\begin{split}
%       &\approxstateArgnt{n} =  \underset{\statey \in \stspace^n}{\text{arg\,min}} \; \mathcal{J}^n(\statey), \\ 
%      &\text{subject to }\;  \approxstateArg{n}{\timeStartArg{n}} =
%\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
%\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
%\end{split}
%\end{equation}
where $\projectorArg{n} : \RR{\fomdim} \rightarrow \stspaceBoundStartArg{n}$ 
is a (spatial) projector onto the start of the $n$th trial space (e.g., $\elltwo$).

>>>>

The \methodAcronym\ framework sequentially computes approximate solutions
$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$, where 
$\approxstateArgnt{n}$ is the solution to the
minimization problems
\begin{equation}\label{eq:tclsrm}
\begin{split}
      &\underset{\statey \in \stspace^n}{\text{minimize}} \; \mathcal{J}^n(\statey), \\
			&\text{subject to } \;  \statey(\timeStartArg{n}) =
\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1}(\timeEndArg{n-1})) & n = 2,\ldots,\nslabs \\
\mathbb{P}^n( \stateFOMIC)& n=1, \end{cases} 
\end{split}
\end{equation}
%\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$ as the solution to the minimization problem,
%\begin{equation}\label{eq:tclsrm}
%\begin{split}
%       &\approxstateArgnt{n} =  \underset{\statey \in \stspace^n}{\text{arg\,min}} \; \mathcal{J}^n(\statey), \\ 
%      &\text{subject to }\;  \approxstateArg{n}{\timeStartArg{n}} =
%\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
%\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
%\end{split}
%\end{equation}
where $\projectorArg{n} : \RR{\fomdim} \rightarrow \stspaceBoundStartArg{n}$ 
is a (spatial) projector onto the start of the $n$th trial space (e.g., $\elltwo$).

### 2019-10-08 section3/space_time.tex
through space--time, proofing

\subsection{\spaceTimeAcronym\ trial spaces}\label{sec:wls_spacetime}
\KTC{Error in first equation. This is the approximate solution not the real
one. Use parallel language to hwo this was done in the last section for
S-reduction trial subspaces. Also make sure the basis is zero at the beginning
of the interval and the affine transformation is taken to be the solution at
the end of the last window. This will not be too trivial, as the solution on
the previous slab enters the trial subspace definition.}

We now briefly consider \spaceTimeAcronym\ trial subspaces. 
The \spaceTimeAcronym\ trial subspace over the $n$th time window approximates
the FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$
with $\approxstateArgnt{n} \in \stspaceSTArg{n}$, where
\begin{equation}\label{eq:st_sttrialspace}
 \stspaceSTArg{n} \defeq
 \Range{\stbasisArgnt{n}} + \stateInterceptArg{n}\otimes \onesFunctionArg{n} .
\end{equation}
Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{0} = \bz$ is the space--time trial basis matrix \textit{function}. At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as,
 \begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{n},
\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

\begin{comment}
each window, we introduce the space--time trial subspace
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
\KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
\KTC{Remove all commas before equations unless it makes sense gramatically.
Add the $n$ superscript to $x_{ref}$}
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
%where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

\end{comment}
\KTC{Next sentence, use parallel language to the last section. It's subtly
different but I think written better there.} Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm}, \methodAcronym\ with 
the \spaceTimeAcronym\ trial subspace computes the sequence of generalized coordinates $\stgenstateArg{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
\begin{align}\label{eq:obj_gen_slab_spacetime}
\begin{split}
&\underset{\stgenstateyArg{n} \in \RR{\stdim}}{\text{minimize}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstateyArg{n} + \stateInterceptArg{n} ), \\ 
      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n}  =
  \begin{cases} 
\mathbb{P}^n(\stbasisArgnt{n-1}(\timeEndArg{n-1})\stgenstateArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
\mathbb{P}^n(\stateFOMIC)
 & n=1. \end{cases}
%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
%\genstateIC & n=1, \end{cases} 
\end{split}
\end{align}
\KTC{In prior equation, remove projection and align properly with ST-LSPG
stuff before.}

%\begin{align}\label{eq:obj_gen_slab_spacetime}
%\begin{split}
%&\stgenstateArg{n} = \underset{\stgenstatey \in \RR{\stdim}}{\text{arg\,min}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n} ), \\ 
%      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n}  =
%  \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMArgnt{1})(0)
% & n=1. \end{cases}
%%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
%%\genstateIC & n=1, \end{cases} 
%\end{split}
%\end{align}

%Inserting the approximation~\eqref{eq:stapprox} into the optimization problem~\eqref{eq:tclsrm}, the
%constrained minimization problem over the $n$th time slab can be re-written as,
%\begin{equation}\label{eq:obj_gen_slab_spacetime} \stgenstate^n =
%\underset{\stgenstatey}{\text{argmin }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \end{equation}
%%\begin{equation}\label{eq:obj_gen_slab} \genstate^n(t) =
%%\underset{\genstatey}{\text{argmin }} \mathcal{J}^n
%%\bigg(\decoder\big(\genstatey(\tau)\big) \bigg) , \end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC
% & n=1. \end{cases} \end{equation}
\subsubsection{Stationary conditions} 
The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial subspaces is as follows: in \spaceTimeAcronym\ trial subspaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial subspaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdim} \otimes \timeSpaceArg{n}$. 
Thus, in the \spaceTimeAcronym\ case the optimization problem is no longer minimizing a functional with respect to a function, but now is minimizing a function with 
respect to vector.  As such, the first-order optimality conditions can be
	derived using standard calculus. Differentiating the objective with respect
	to $\stgenstatey$ and setting equal to zero yields \KTC{Make sure that xref
	is replaced with the last solution at the previous time step}
\begin{equation}\label{eq:st_stationary}
 \intSlabArg{n} \bigg[ \stbasisDotArg{n}{t}^T  - \stbasisArg{n}{t}^T \bigg[\frac{\partial
\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
\stateInterceptArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{n},t) \bigg) dt = \bz.\end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC &
%n=1. \end{cases} \end{equation}
For \spaceTimeAcronym\ trial subspaces, the stationary conditions comprise an algebraic system of integral equations, as opposed to a system of differential equations.

>>>>

\subsection{\spaceTimeAcronym\ trial spaces}\label{sec:wls_spacetime}
%\KTC{Error in first equation. This is the approximate solution not the real
%one. Use parallel language to hwo this was done in the last section for
%S-reduction trial subspaces. Also make sure the basis is zero at the beginning
%of the interval and the affine transformation is taken to be the solution at
%the end of the last window. This will not be too trivial, as the solution on
%the previous slab enters the trial subspace definition.}

%We now briefly consider \spaceTimeAcronym\ trial subspaces. 
The \spaceTimeAcronym\ trial subspace over the $n$th time window approximates
the FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$
with $\approxstateArgnt{n} \in \stspaceSTArg{n}$, where
\begin{equation}\label{eq:st_sttrialspace}
 \stspaceSTArg{n} \defeq
 \Range{\stbasisArgnt{n}} + \stateInterceptSTArg{n}\otimes \onesFunctionArg{n} .
\end{equation}
Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{0} = \bz$ is the space--time trial basis matrix \textit{function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$ provides the affine transformation. 
For \spaceTimeAcronym\ trial subspaces, we take the affine transformation to be $\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for $n=2,\ldots,\nslabs$ and $\stateInterceptSTArg{1} = \stateFOMIC$; this definition of the trial subspace automatically satisfies the boundary conditions.
At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as,
 \begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},
\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window. 
\begin{comment}
each window, we introduce the space--time trial subspace
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
\KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptSTArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
\KTC{Remove all commas before equations unless it makes sense gramatically.
Add the $n$ superscript to $x_{ref}$}
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
%where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

\end{comment}
Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm} implies that \methodAcronym\ with \spaceTimeAcronym\ trial subspaces sequentially computes solutions $\stgenstateArg{n}$, $n = 1,\ldots,\nslabs$ that satisfy
\begin{align}\label{eq:obj_gen_slab_spacetime}
\begin{split}
&\underset{\stgenstateyArg{n} \in \RR{\stdim}}{\text{minimize}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstateyArg{n} + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n} ).% \\ 
%      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n}  =0
%  \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}(\timeEndArg{n-1})\stgenstateArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMIC)
% & n=1. \end{cases}
%%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
%\genstateIC & n=1, \end{cases} 
\end{split}
\end{align}
%\begin{align}\label{eq:obj_gen_slab_spacetime}
%\begin{split}
%&\stgenstateArg{n} = \underset{\stgenstatey \in \RR{\stdim}}{\text{arg\,min}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n} ), \\ 
%      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n}  =
%  \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMArgnt{1})(0)
% & n=1. \end{cases}
%%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
%%\genstateIC & n=1, \end{cases} 
%\end{split}
%\end{align}

%Inserting the approximation~\eqref{eq:stapprox} into the optimization problem~\eqref{eq:tclsrm}, the
%constrained minimization problem over the $n$th time slab can be re-written as,
%\begin{equation}\label{eq:obj_gen_slab_spacetime} \stgenstate^n =
%\underset{\stgenstatey}{\text{argmin }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \end{equation}
%%\begin{equation}\label{eq:obj_gen_slab} \genstate^n(t) =
%%\underset{\genstatey}{\text{argmin }} \mathcal{J}^n
%%\bigg(\decoder\big(\genstatey(\tau)\big) \bigg) , \end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC
% & n=1. \end{cases} \end{equation}
\subsubsection{Stationary conditions} 
The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial subspaces is as follows: in \spaceTimeAcronym\ trial subspaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial subspaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdim} \otimes \timeSpaceArg{n}$. 
Thus, in the \spaceTimeAcronym\ case the optimization problem is no longer minimizing a functional with respect to a function, but now is minimizing a function with 
respect to vector.  As such, the first-order optimality conditions can be
	derived using standard calculus. Differentiating the objective with respect
	to $\stgenstatey$ and setting equal to zero yields 
\begin{equation}\label{eq:st_stationary}
 \intSlabArg{n} \bigg[ \stbasisDotArg{n}{t}^T  - \stbasisArg{n}{t}^T \bigg[\frac{\partial
\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
\stateInterceptSTArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},t) \bigg) dt = \bz.\end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC &
%n=1. \end{cases} \end{equation}
For \spaceTimeAcronym\ trial subspaces, the stationary conditions comprise an algebraic system of integral equations, as opposed to a system of differential equations.

### 2019-10-08 section3/tclsrm_formulation.tex
through space--time, proofing

\subsection{\methodAcronym\ summary}
This section outlined the \methodAcronym\ method for model reduction. In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial subspace for a series of time windows: \methodAcronym\  
\KTC{Here, use the same language as earlier regarding sequential computation
of these coordinates that satisfy...} computes the generalized coordinates
$\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the
minimization problems, \KTC{remove superscript from optimizaton variable
everywhere}
\begin{equation*}
\underset{\stateyArgnt{} \in \stspace^n}{\text{minimize } }
\mathcal{J}^n(\stateyArgnt{}), \qquad n = 1,\ldots,\nslabs,
\end{equation*}
where,
\begin{align*}
\mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArg{n} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
\RR{},
\end{align*}
subject to the boundary and continuity conditions. For the \spatialAcronym\ trial subspaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
\begin{align*}\
&\massArg{n}  \genstateDotArg{n}{t}  -  \basisspaceTArg{n}
\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \massArg{n} \adjointArg{n}{t} , \\
 &\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial
\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspaceArg{n}
 \adjointArg{n}{t} = \nonumber \\ 
&-\basisspaceTArg{n} \big[
\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \bigg( \mathbf{I} -
\basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArgt{n}{t} \bigg)
 \bigg( \basisspace \genstateDotArg{n}{t}  -
\velocity(\veloargsromn) \bigg),  &
%\genstateArg{n}{\timeStartArg{n}} = \begin{cases}
%\genstate^{n-1}(\timeEndArg{n-1}) & n=2,\ldots,\nslabs, \\
%\basisspace^T(\approxstate_0 - \stateIntercept) & n=1, \end{cases}\\ &
%\adjointArg{n}{\timeEndArg{n}} = \boldsymbol 0.
\end{align*} 
subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}. The Euler-Lagrange equations expose the fact that, for \spatialAcronym\
trial subspaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
controller that enforces the residual minimization property.

Section~\ref{sec:wls_spacetime} additionally considered \spaceTimeAcronym\ trial subspaces. For \spaceTimeAcronym\ trial subspaces, in where the generalized coordinates are no longer functions, the stationary conditions correspond to an algebraic system of equations.

>>>>

\subsection{\methodAcronym\ summary}
This section outlined the \methodAcronym\ method for model reduction. In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial subspace for a series of time windows; i.e.,  
%\methodAcronym\  
%computes the generalized coordinates
%$\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the
%minimization problems, 
\methodAcronym\ sequentially computes approximate solutions
$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$, where 
$\approxstateArgnt{n}$ is the solution to the
minimization problem

\begin{equation*}
\underset{\stateyArgnt{} \in \stspace^n}{\text{minimize } }
\mathcal{J}^n(\stateyArgnt{}), \qquad n = 1,\ldots,\nslabs,
\end{equation*}
where,
\begin{align*}
\mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArg{n} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
\RR{},
\end{align*}
subject to the boundary conditions. For the \spatialAcronym\ trial subspaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
\begin{align*}\
&\massArg{n}  \genstateDotArg{n}{t}  -  \basisspaceTArg{n}
\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \massArg{n} \adjointArg{n}{t} , \\
 &\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial
\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspaceArg{n}
 \adjointArg{n}{t} = \nonumber \\ 
&-\basisspaceTArg{n} \big[
\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \bigg( \mathbf{I} -
\basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArgt{n}{t} \bigg)
 \bigg( \basisspace \genstateDotArg{n}{t}  -
\velocity(\veloargsromn) \bigg),  &
%\genstateArg{n}{\timeStartArg{n}} = \begin{cases}
%\genstate^{n-1}(\timeEndArg{n-1}) & n=2,\ldots,\nslabs, \\
%\basisspace^T(\approxstate_0 - \stateIntercept) & n=1, \end{cases}\\ &
%\adjointArg{n}{\timeEndArg{n}} = \boldsymbol 0.
\end{align*} 
subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}. The Euler--Lagrange equations expose the fact that, for \spatialAcronym\
trial subspaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
controller that enforces the residual minimization property. 
Section~\ref{sec:wls_spacetime} additionally considered \spaceTimeAcronym\ trial subspaces. For \spaceTimeAcronym\ trial subspaces, in where the generalized coordinates are no longer functions, the stationary conditions correspond to an algebraic system of equations.
